Mining Scenario/Transaction Pairs in HW/SW Co-verification Using Symbolic Execution

### 2017-04-04 paper.tex
section 3.1

\textbf{Step 2: Constructing Co-verification Model}
An unified co-verification model is constructed by sequentializing 
HW and its interacting FW threads, thereby generating a sequential 
single-threaded co-verification model in C.  The motivation behind
sequentialization is based on our observation that many practical industrial
co-designs exhibit producer-consumer interaction behavior, such as IBM 
Text Accelerator co-design~\cite{polig2014micro,polig2014fpl}, 
IBM coherent FFT coprocessor~\cite{giefers2015accelerating}, RockBox 
Media Player~\cite{hvc}, linux device driver interacting with QEMU emulated
hardware code~\cite{codes14} and others.  We observe that in practise, a FW 
thread is mostly independent of the HW thread they interact with~\cite{hvc,codes14}.  
Specifically, a FW thread is only responsible for configuring the 
HW transactions by writing to memory mapped registers or polling 
the interrupt status register for data transmission or receiving 
incoming data packets. This interaction pattern exhibits producer-consumer 
relationship, which is also evident in case of our UART example in 
Figure~\ref{fig:interleaving1}.  
%
We now formally define a single-threaded sequential co-verification model, 
denoted by $\mathcal{M}_{seq}$, obtained by sequentializing the HW and FW 
models.
%
\[ 
\begin{array}[t]{@{}lll}
  \mathcal{M}_{seq} & {:}{:}{=} & \mathit{Procedure} \\
\mathit{Procedure} & {:}{:}{=} & \mathit{Statement} \mid \mathit{Procedure} \\
\mathit{Statement} & {:}{:}{=} & CStatement \mid \mathit{moduleName(Var_1,\dots,Var_n)} \\
\mathit{CStatement} & {:}{:}{=} & x {:}{=} exp \mid \mathit{ITE}(b, s_1,s_2) \mid \mathit{s_1;s_2} \mid loop\{s\} \\ 
\end{array}
\]
%
Consider sets of expressions $Exp$ and Boolean expressions $BExp$ 
over variables $Var$ of $\mathcal{M}_{seq}$ such that $(exp \in Exp)$ and 
$(b \in BExp)$. A $\mathit{CStatement}$ is an assignment,
conditional, sequential concatenation or a loop.
$\mathit{moduleName(Var_1,\dots,Var_n)}$ invokes the top level 
procedure of the $\mathcal{SN}$. Recall that $\mathcal{SN}$ is a
software netlist representation of HW RTL models.
The interface variables $\{Var\_1,\dots,Var\_n\}$ 
(input, output and inout ports) are declared inside the 
FW to communicate with $\mathcal{SN}$. \\ 
%

\textbf{Step 3: Sequentialization based on Producer-Consumer Interaction}
%
A communicating pair of executions of 
$\mathcal{FT}$ and $\mathcal{HT}$ is an {\em interaction}, which 
takes place through shared variables (or shared registers)
only.  An interaction is said to exhibit {\em producer-consumer}
interaction pattern if 1) the set of variables $V$ written by 
the execution of $\mathcal{FT}$ and and the set of variables 
$V'$ written by the execution of $\mathcal{HT}$ are disjoint 
($V \cap V' = \emptyset$), (2) and the executions 
of $\mathcal{FT}$ and $\mathcal{HT}$ consumes values 
in $V'$ and $V$, respectively.
%
%
\begin{figure}[t]
\scalebox{.55}{\import{figures/}{interleaving.pspdftex}}
\caption{Producer Consumer Interactions Pattern
\label{fig:interleaving1}}
\end{figure}
%

\textbf{Step 4: Property Driven Slicing of Co-verification Model}
A property-driven slicing is performed on the unified model, $\mathcal{M}_{seq}$.  
This step is purely syntactic, meaning that we perform a backward dependency 
analysis starting from the property which only preserve those 
program fragments that are relevant to a given property. The sliced program 
is then passed to the symbolic execution engine for path-based exploration. \\
%
%=======================================================================================
%\subsection{\verifox: Path-wise Symbolic Execution Tool} 
%=======================================================================================
%\Omit{
%\begin{figure}[tp]
%\centering
%\vspace*{0.3cm}
%\scalebox{.50}{\import{figures/}{work-flow.pspdftex}}
%\caption{Automated Slicing and Trace partitioning\label{fig:desc}}
%\end{figure}
%}
\Omit{
\begin{figure}[t]
\tikzstyle{rblock}=[rounded corners,draw=black]
\tikzstyle{block}=[draw=black]
\begin{tikzpicture}

\node (v2c) [rblock] {\scs{\emph{v2c}}};
\node (snl) [block,below of=v2c] {\scs{Software-netlist}};
\draw[->] (v2c) -- (snl);

\node (HD) [above left=.3cm and .3cm of v2c] {\scs{\parbox{8em}{Hardware Design \\ (in Verilog)}}};
\draw[->] (HD) -- (v2c);

\node (prop) [above right=.3cm and .3cm of v2c] {\scs{\parbox{8em}{Property \\ (in SVA)}}};
\draw[->] (prop) -- (v2c);

\node (cmm) [block,below of=snl] {\scs{\parbox{9em}{Composite Monolithic Model}}};
\draw[->] (snl) -- (cmm);

\node (dcmm) [left=1cm of cmm] {};
\node (fw) [above=.1cm of dcmm  ] {\scs{Firmware}};
\node (as) [below=.1cm of dcmm] {\scs{Assumptions}};
\draw[->] (fw) -- (cmm);
\draw[->] (as) -- (cmm);

\node (pds) [block,below of=cmm] {\scs{\parbox{8em}{Property-driven slicing}}};
\draw[->] (cmm) -- (pds);

\node (tp) [block,below=.8cm of pds] {\scs{Trace partitioning}};
\node (iss) [block,below right=.5cm  and .05cm of tp.east] {\scs{\parbox{6em}{Incremental SAT solving}}};
\node (sat) [rblock,below=1.5cm of tp] {\scs{SAT solver}};
\node (ipp) [block,left=2cm of iss] {\scs{\parbox{6em}{Infeasible path pruning}}};

\draw [<->] (tp) -- (iss);
\draw [<->] (tp) -- (ipp);
\draw [<->] (sat) -- (iss);
\draw [<->] (sat) -- (ipp);

\node (vfbox) [rblock,fit={(tp) (iss) (sat) (ipp)}] {};
\draw[->] (pds) -- (vfbox);

\node (verifox) [right=.2cm of tp] {\scs{{\bf \verifox}}};

\node (safe) [above right=.2cm and .2cm of vfbox.east] {\scs{Verification Successful}};
\node (unsafe) [below right=.2cm and .2cm of vfbox.east] {\scs{\parbox{8em}{Verification Failed (Counterexample)}}};

\draw[->] (vfbox.9) -- (safe);
\draw[->] (vfbox.-12) -- (unsafe);

\end{tikzpicture}
\caption{Automated Slicing and Trace partitioning\label{fig:desc}}
\end{figure}
}
%
\input{verifox}
%
\subsection{Approach 2: Monolithic Co-verification Technique}
%
\begin{figure}[t]
\begin{center}
\includegraphics[scale=0.6]{figures/traditional_flow.pdf}
\caption{Conventional HW/SW Co-verification Flow}
\label{fig:conventional}
\end{center}
\end{figure}

Figure~\ref{fig:conventional} illustrates the conventional approach to HW/SW
co-veri\-fi\-cation, as implemented in the state-of-the-art co-verification
tool, \hwcbmcv~\cite{CKY03,vlsid13}.  \hwcbmcv is the only open-source
formal co-verification tool that supports Verilog RTL (System Verilog 2009
standard) and C/C++ firmware.  In contrast to our proposed approach,
\hwcbmcv maintains two separate flows for HW and FW.  The top flow in
Figure~\ref{fig:conventional} uses synthesis to obtain either a bit-level or
a word-level netlist from Verilog RTL.  The bottom flow illustrates the
translation of the C program into static single assignment (SSA) form. 
These two flows meet only at the solver.  Thus, \hwcbmcv generates a
monolithic formula from the C and RTL description, which is then checked
with SAT/SMT solvers.  \hwcbmcv provides specific handshake primitives such
as $next\_timeframe()$ and $set\_inputs()$ to model FW-HW communications.
\rmcmt{
To this end, \hwcbmcv constructs a co-verification model through 
in tandem symbolic execution of the HW and FW models.  A bit-level 
netlist representation in And Inverter Graph (AIG) format or word-level netlist in
format similar to SMT-LIB standards is synthesized from a HW Register Transfer 
Level (RTL) implementation. Similarly, the SW, given in C, is automatically 
translated into a Static Single Assignment (SSA) form.  A monolithic verification 
condition is generated from these two models which is then solved using a 
}
%
\Omit{
The hardware and firmware run independently of each other. 
The communication between them takes place through these function calls. 
}
%===============================================================================
\section{Experimental Results}
%===============================================================================
%
We report experimental results for FW-HW co-verification of a UART IP and a
SoC design.  All our experiments were performed on an Intel Xeon 3.07\,GHz
machine with 48\,GB RAM.  All times reported are in seconds.  The timeout 
for all our experiments was set to 2~hours.  
%
%-------------------------------------------------------------------------------
\subsection{FW-HW co-verification of UART}
%-------------------------------------------------------------------------------

@@

>>>>

\textbf{Step 2: Sequentialization based on Producer-Consumer Interaction}
%
A communicating pair of executions of 
$\mathcal{FT}$ and $\mathcal{HT}$ is an {\em interaction}, which 
takes place through shared variables (or shared registers)
only.  An interaction is said to exhibit {\em producer-consumer}
interaction pattern if 1) the set of variables $V$ written by 
the execution of $\mathcal{FT}$ and and the set of variables 
$V'$ written by the execution of $\mathcal{HT}$ are disjoint 
($V \cap V' = \emptyset$), (2) and the executions 
of $\mathcal{FT}$ and $\mathcal{HT}$ consumes values 
in $V'$ and $V$, respectively.
%
Figure~\ref{fig:interleaving1} illustrates the FW-HW interaction pattern 
for the UART IP. The FW CFG is shown in the left and the 
HW RTL code snippet is shown in the right. We abuse notations and 
represent $\mathcal{HT}$ in the form of RTL and $\mathcal{FT}$ as CFG
for our illustration purpose. Here, $\mathcal{HT}$ is the producer of interrupt 
status signal, $irq\_o$, which is read by $\mathcal{FT}$ in state $s_1$. 
The signal $irq\_o$ is  triggered when transmitter buffer is 
{\em EMPTY} or receiver buffer is {\em FULL}. 
$\mathcal{HT}$ also produces interrupt status by writing to the 
memory mapped register, $UART\_IS$, which is consumed by 
$\mathcal{FT}$ in state $s_2$.  In another interaction, $\mathcal{HT}$
sends a data-packet to $UART\_TR$ which is consumed by $\mathcal{FT}$ in state $s_7$.
Similarly, $\mathcal{FT}$ is the producer of $data$, $(outb(data,UART\_TR))$, 
in state $s_6$ which is consumed by $\mathcal{HT}$. These interaction patterns 
between $\mathcal{HT}$ and $\mathcal{FT}$ is exploited to generate a 
sequential unified {\em program netlist} model which is defined next.\\ \\
%
\begin{figure}[t]
\scalebox{.55}{\import{figures/}{interleaving.pspdftex}}
\caption{Producer Consumer Interactions Pattern
\label{fig:interleaving1}}
\end{figure}
%

\textbf{Step 3: Constructing Co-verification Model}
An unified co-verification model is constructed by sequentializing 
HW and its interacting FW threads, thereby generating a sequential 
single-threaded co-verification model in C.  The motivation behind
sequentialization is based on our observation that many practical industrial
co-designs exhibit producer-consumer interaction behavior, such as IBM 
Text Accelerator co-design~\cite{polig2014micro,polig2014fpl}, 
IBM coherent FFT coprocessor~\cite{giefers2015accelerating}, RockBox 
Media Player~\cite{hvc}, linux device driver interacting with QEMU emulated
hardware code~\cite{codes14} and others.  We observe that in practise, a FW 
thread is mostly independent of the HW thread they interact with~\cite{hvc,codes14}.  
Specifically, a FW thread is only responsible for configuring the 
HW transactions by writing to memory mapped registers or polling 
the interrupt status register for data transmission or receiving 
incoming data packets. This interaction pattern exhibits producer-consumer 
relationship, which is also evident in case of our UART example in 
Figure~\ref{fig:interleaving1}.  
%
We now formally define a single-threaded sequential co-verification model, 
denoted by $\mathcal{M}_{seq}$, obtained by sequentializing the HW and FW 
models.
%
\[ 
\begin{array}[t]{@{}lll}
  \mathcal{M}_{seq} & {:}{:}{=} & \mathit{Procedure} \\
\mathit{Procedure} & {:}{:}{=} & \mathit{Statement} \mid \mathit{Procedure} \\
\mathit{Statement} & {:}{:}{=} & CStatement \mid \mathit{moduleName(Var_1,\dots,Var_n)} \\
\mathit{CStatement} & {:}{:}{=} & x {:}{=} exp \mid \mathit{ITE}(b, s_1,s_2) \mid \mathit{s_1;s_2} \mid loop\{s\} \\ 
\end{array}
\]
%
Consider sets of expressions $Exp$ and Boolean expressions $BExp$ 
over variables $Var$ of $\mathcal{M}_{seq}$ such that $(exp \in Exp)$ and 
$(b \in BExp)$. A $\mathit{CStatement}$ is an assignment,
conditional, sequential concatenation or a loop.
$\mathit{moduleName(Var_1,\dots,Var_n)}$ invokes the top level 
procedure of the $\mathcal{SN}$. Recall that $\mathcal{SN}$ is a
software netlist representation of HW RTL models.
The interface variables $\{Var\_1,\dots,Var\_n\}$ 
(input, output and inout ports) are declared inside the 
FW to communicate with $\mathcal{SN}$. \\ 
%

\textbf{Step 4: Property Driven Slicing of Co-verification Model}
A property-driven slicing is performed on the unified model, $\mathcal{M}_{seq}$.  
This step is purely syntactic, meaning that we perform a backward dependency 
analysis starting from the property which only preserve those 
program fragments that are relevant to a given property. The sliced program 
is then passed to the symbolic execution engine for path-based exploration. \\
%
%=======================================================================================
%\subsection{\verifox: Path-wise Symbolic Execution Tool} 
%=======================================================================================
\input{verifox}
%
\subsection{Approach 2: Monolithic Co-verification Technique}
%
\begin{figure}[t]
\begin{center}
\includegraphics[scale=0.6]{figures/traditional_flow.pdf}
\caption{Conventional HW/SW Co-verification Flow}
\label{fig:conventional}
\end{center}
\end{figure}
%
Figure~\ref{fig:conventional} illustrates the conventional approach to HW/SW
co-veri\-fi\-cation, as implemented in the state-of-the-art co-verification
tool, \hwcbmcv~\cite{CKY03,vlsid13}.  \hwcbmcv is the only open-source
formal co-verification tool that supports Verilog RTL (System Verilog 2009
standard) and C/C++ firmware.  In contrast to our proposed approach,
\hwcbmcv maintains two separate flows for HW and FW.  The top flow in
Figure~\ref{fig:conventional} uses synthesis to obtain either a bit-level or
a word-level netlist from Verilog RTL.  The bottom flow illustrates the
translation of the C program into static single assignment (SSA) form. 
These two flows meet only at the solver.  Thus, \hwcbmcv generates a
monolithic formula from the C and RTL description, which is then checked
with SAT/SMT solvers.  \hwcbmcv provides specific handshake primitives such
as $next\_timeframe()$ and $set\_inputs()$ to model FW-HW communications.
\rmcmt{
To this end, \hwcbmcv constructs a co-verification model through 
in tandem symbolic execution of the HW and FW models.  A bit-level 
netlist representation in And Inverter Graph (AIG) format or word-level netlist in
format similar to SMT-LIB standards is synthesized from a HW Register Transfer 
Level (RTL) implementation. Similarly, the SW, given in C, is automatically 
translated into a Static Single Assignment (SSA) form.  A monolithic verification 
condition is generated from these two models which is then solved using a 
}
%
\Omit{
The hardware and firmware run independently of each other. 
The communication between them takes place through these function calls. 
}
%===============================================================================
\section{Experimental Results}
%===============================================================================
%
We report experimental results for FW-HW co-verification of a UART IP and a
SoC design.  All our experiments were performed on an Intel Xeon 3.07\,GHz
machine with 48\,GB RAM.  All times reported are in seconds.  The timeout 
for all our experiments was set to 2~hours.  
%

\textit{\textbf{State-of-the-art Co-verification tools:}} 
%There exists plethora of tools for either SW verification or HW verification.
%
Despite extensive use of model checking and other formal methods in 
SW verification or HW verification domain, building automated 
co-verification tools using formal methods has received 
little attention in the past.  We are not aware of any automated 
formal co-verification tool in the literature that can readily accept 
co-designs written in C and Verilog RTL. In this paper, we address this gap. 
%
%-------------------------------------------------------------------------------
\subsection{FW-HW co-verification of UART}
%-------------------------------------------------------------------------------

@@

% ===============================================================================
\section{Limitations of Proposed Approach}
% ===============================================================================
